Crystal as a thermodynamic computer

### 2024-09-10 paper/main.tex
modify main figure, rewrited partD, still some work need to do

    %Shape: Axis 2D [id:dp47589449610201395] 
    % \draw  (252,306) -- (464.6,306)(252,45.4) -- (252,306) -- cycle (457.6,301) -- (464.6,306) -- (457.6,311) (247,52.4) -- (252,45.4) -- (257,52.4)  ;
    \draw[->][line width=1.5pt] (257,315) -- (464.6,315) node[anchor=north west] {};
    % 绘制 Y 轴
    \draw[->][line width=1.5pt] (257,315) -- (257,45.4) node[anchor=south east] {};
    % \node at (358.3, 306) [anchor=north] {$Time$};
    \draw (358.3,320) node [anchor=north][inner sep=0.75pt]   [align=left] {${Time}$};
    % \node at (252, 175.7) [anchor=east] {$Space$};
    \draw (215,175.7) node [anchor=west][inner sep=0.75pt]   [align=left] {${Space}$};
    %Rounded Rect [id:dp5816101232993451] 
    \path  [shading=_qxg3a7xt1,_ue4dovf1e,path fading= _htcz8osfa ,fading transform={xshift=2}] (281.6,76.28) .. controls (281.6,66.96) and (289.16,59.4) .. (298.48,59.4) -- (349.12,59.4) .. controls (358.44,59.4) and (366,66.96) .. (366,76.28) -- (366,274.52) .. controls (366,283.84) and (358.44,291.4) .. (349.12,291.4) -- (298.48,291.4) .. controls (289.16,291.4) and (281.6,283.84) .. (281.6,274.52) -- cycle ; % for fading 
     % \draw   (281.6,76.28) .. controls (281.6,66.96) and (289.16,59.4) .. (298.48,59.4) -- (349.12,59.4) .. controls (358.44,59.4) and (366,66.96) .. (366,76.28) -- (366,274.52) .. controls (366,283.84) and (358.44,291.4) .. (349.12,291.4) -- (298.48,291.4) .. controls (289.16,291.4) and (281.6,283.84) .. (281.6,274.52) -- cycle ; % for border

@@

    % Text Node
    \draw (379,111.8) node [anchor=north west][inner sep=0.75pt]   [align=left] {$\displaystyle \cdots $};
    % Text Node
    \draw (379,141.8) node [anchor=north west][inner sep=0.75pt]   [align=left] {$\displaystyle \cdots $};
    % Text Node
    \draw (379,172.09) node [anchor=north west][inner sep=0.75pt]   [align=left] {$\displaystyle \cdots $};
    % Text Node
    \draw (379,241.8) node [anchor=north west][inner sep=0.75pt]   [align=left] {$\displaystyle \cdots $};
    % Text Node
    \draw (283,193.97) node [anchor=north west][inner sep=0.75pt]   [align=left] {$\displaystyle \vdots $};
    % Text Node
    \draw (322.5,193.97) node [anchor=north west][inner sep=0.75pt]   [align=left] {$\displaystyle \vdots $};
    % Text Node
    \draw (363,193.97) node [anchor=north west][inner sep=0.75pt]   [align=left] {$\displaystyle \vdots $};
    % Text Node
    \draw (420.1,194) node [anchor=north west][inner sep=0.75pt]   [align=left] {$\displaystyle \vdots $};
    % Text Node
    \draw (382.4,194.52) node [anchor=north west][inner sep=0.75pt]   [align=left] {$\displaystyle \ddots $};
    % Text Node
    \draw (292,45) node [anchor=north west][inner sep=0.75pt]   [align=left] {${Heat\ Zone}$};
    % Text Node
    %\draw (220.87,45.67) node [anchor=north west][inner sep=3pt]   [align=left] {\large {$\displaystyle \mathtt{(a)}$}};
    % Text Node
    %\draw (490.67,121.47) node [anchor=north west][inner sep=3pt]   [align=left] {\large {$\displaystyle \mathtt{(b)}$}};
    % Text Node
    \draw (492.00,150.41) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\tiny {$p$}}};
    % Text Node
    \draw (479.75,168.95) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\tiny {$q$}}};
    % Text Node
    \draw (492.45,188.12) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\tiny {$r$}}};
    % Text Node
    \draw (525.00,166.95) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\tiny {$q'$}}};
    % Text Node
    \draw (504.00,168.45) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\tiny {$A$}}};

    \draw [<->] [line width=1.0] (283.2, 295.4) -- (365.2, 295.4) ;
    \draw [->] [line width=1.0] (310, 73) -- (337, 73);
    % \draw   (283.2,286.4) .. controls (283.2,291.07) and (285.53,293.4) .. (290.2,293.4) -- (314.2,293.4) .. controls (320.87,293.4) and (324.2,295.73) .. (324.2,300.4) .. controls (324.2,295.73) and (327.53,293.4) .. (334.2,293.4)(331.2,293.4) -- (358.2,293.4) .. controls (362.87,293.4) and (365.2,291.07) .. (365.2,286.4) ;    \draw   [->] (312.63,73.33) -- (336.63,73.33) ;
    % Text Node
    \draw (319.33,63.67) node [anchor=north west][inner sep=0.75pt]   [align=left] {$\displaystyle v$};
    % Text Node
    \draw (306.20,298.47) node [anchor=north west][inner sep=0.75pt]   [align=left] {$\displaystyle Width$};

    \end{tikzpicture}
    \caption{A surface programmable thermodynamic computer (TC) that implements the rule 110 elementary cellular automata. The vertical direction is the space axis,
    while the horizontal direction is the time axis.
    The computation is done by moving the heat zone (orange colored area) from the input surface (left) to the output surface (right) slowly with a constant speed $v$.
    The TC has a translational invariant structure, and each unit cell (inset) is a $110$ rule cellular automata gadget that consists of 5 mutually correlated spins (white circles).
    \jinguo{please add symbols for space, time and width}}
    \label{Surface-programmable-thermodynamic computer}
\end{figure}

@@

\subsection{Simulated Annealing}

@@

\subsection{Cooling computation with a moving heat bath}\label{cooling-computation-with-moving-heat-bath}

@@

The computational process of the CA is divided into three steps: determining the state of each input cell, iteratively applying the transition rules to generate the next generation of cells, and identifying a certain generation of cells as the output. In fact, the width of the time axis in the Ising TC corresponds to the number of iterations. Hence, the smallest layer of spins on the time axis should be the input layer, and the largest layer of spins should be the output layer.

Given that each ground state of the TC corresponds to a valid computational process of the CA for a given input, the computational process of the TC is essentially reduced to finding the ground state of the TC with the spin configuration in the input layers fixed.

We focus here on the thermodynamics-driven approach to finding ground states, specifically the heating-annealing method. This involves gradually increasing and then decreasing the temperature of the heat bath with which the TC interacts. The TC may employ heat fluctuations to explore a state space in the first stage and stabilize on ground state during the second stage as long as the annealing stage is long enough. However, it is convinced that this stage can last for exponential time for some hard spin-glass problems.

Due to the structure of the TC, we can allow only a portion of the system, rather than the entire system, to interact with the heat bath, while ignoring the interaction energy between the boundary spins and the exterior. After completing the heating-annealing procedure on one part of the system, we move the heat bath to another part and repeat the same process. This approach effectively finds a local ground state for each part of the system, which is reasonable when the global ground state can be divided into multiple local components.

More specifically, the computation, progressing from the input layer to the output layer, involves the following steps: (1) Initialize the input layer configuration by adjusting the onsite energy for each spin. (2) Determine the number of layers $W$ that will interact with the external heat bath in each heating-annealing process. Group the layers in sets of $W$ from input to output, labeling them as groups $1, 2, \cdots, \frac{D}{W}$. The number of heating-annealing repetitions will then be $\frac{D}{W}$, where D is the total number of layers. (3) Perform the heating-annealing process on each group sequentially.

However, solving the ground state from the non-deterministic direction is at least as hard as solving the circuit satisfiability problem, which is NP-complete~\cite{Moore2011}.

>>>>

    %Shape: Axis 2D [id:dp47589449610201395] 
    % \draw  (252,306) -- (464.6,306)(252,45.4) -- (252,306) -- cycle (457.6,301) -- (464.6,306) -- (457.6,311) (247,52.4) -- (252,45.4) -- (257,52.4)  ;
    \draw[->][line width=1.5pt] (257,315) -- (464.6,315) node[anchor=north west] {};
    % 绘制 Y 轴
    \draw[->][line width=1.5pt] (257,315) -- (257,45.4) node[anchor=south east] {};
    % \node at (358.3, 306) [anchor=north] {$Time$};
    % \draw (358.3,320) node [anchor=north][inner sep=0.75pt]   [align=left] {${Time}$};
    % \node at (252, 175.7) [anchor=east] {$Space$};
    % \draw (215,175.7) node [anchor=west][inner sep=0.75pt]   [align=left] {${Space}$};
    %Rounded Rect [id:dp5816101232993451] 
    \path  [shading=_qxg3a7xt1,_ue4dovf1e,path fading= _htcz8osfa ,fading transform={xshift=2}] (281.6,76.28) .. controls (281.6,66.96) and (289.16,59.4) .. (298.48,59.4) -- (349.12,59.4) .. controls (358.44,59.4) and (366,66.96) .. (366,76.28) -- (366,274.52) .. controls (366,283.84) and (358.44,291.4) .. (349.12,291.4) -- (298.48,291.4) .. controls (289.16,291.4) and (281.6,283.84) .. (281.6,274.52) -- cycle ; % for fading 
     % \draw   (281.6,76.28) .. controls (281.6,66.96) and (289.16,59.4) .. (298.48,59.4) -- (349.12,59.4) .. controls (358.44,59.4) and (366,66.96) .. (366,76.28) -- (366,274.52) .. controls (366,283.84) and (358.44,291.4) .. (349.12,291.4) -- (298.48,291.4) .. controls (289.16,291.4) and (281.6,283.84) .. (281.6,274.52) -- cycle ; % for border

@@

    % Text Node
    \draw (379,111.8) node [anchor=north west][inner sep=0.75pt]   [align=left] {$\displaystyle \cdots $};
    % Text Node
    \draw (379,141.8) node [anchor=north west][inner sep=0.75pt]   [align=left] {$\displaystyle \cdots $};
    % Text Node
    \draw (379,172.09) node [anchor=north west][inner sep=0.75pt]   [align=left] {$\displaystyle \cdots $};
    % Text Node
    \draw (379,241.8) node [anchor=north west][inner sep=0.75pt]   [align=left] {$\displaystyle \cdots $};
    % Text Node
    \draw (283,193.97) node [anchor=north west][inner sep=0.75pt]   [align=left] {$\displaystyle \vdots $};
    % Text Node
    \draw (322.5,193.97) node [anchor=north west][inner sep=0.75pt]   [align=left] {$\displaystyle \vdots $};
    % Text Node
    \draw (363,193.97) node [anchor=north west][inner sep=0.75pt]   [align=left] {$\displaystyle \vdots $};
    % Text Node
    \draw (420.1,194) node [anchor=north west][inner sep=0.75pt]   [align=left] {$\displaystyle \vdots $};
    % Text Node
    \draw (382.4,194.52) node [anchor=north west][inner sep=0.75pt]   [align=left] {$\displaystyle \ddots $};
    % Text Node
    \draw (292,45) node [anchor=north west][inner sep=0.75pt]   [align=left] {${Heat\ Zone}$};
    % Text Node
    \draw (492.00,150.41) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\tiny {$p$}}};
    % Text Node
    \draw (479.75,168.95) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\tiny {$q$}}};
    % Text Node
    \draw (492.45,188.12) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\tiny {$r$}}};
    % Text Node
    \draw (525.00,166.95) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\tiny {$q'$}}};
    % Text Node
    \draw (504.00,168.45) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\tiny {$A$}}};

    \draw [<->] [line width=1.0] (283.2, 295.4) -- (365.2, 295.4) ;
    \draw [->] [line width=1.0] (310, 73) -- (337, 73);
    % \draw   (283.2,286.4) .. controls (283.2,291.07) and (285.53,293.4) .. (290.2,293.4) -- (314.2,293.4) .. controls (320.87,293.4) and (324.2,295.73) .. (324.2,300.4) .. controls (324.2,295.73) and (327.53,293.4) .. (334.2,293.4)(331.2,293.4) -- (358.2,293.4) .. controls (362.87,293.4) and (365.2,291.07) .. (365.2,286.4) ;    \draw   [->] (312.63,73.33) -- (336.63,73.33) ;
    % Text Node
    \draw (319.33,63.67) node [anchor=north west][inner sep=0.75pt]   [align=left] {$\displaystyle v$};
    % Text Node
    \draw (317.20,298.47) node [anchor=north west][inner sep=0.75pt]   [align=left] {$W$};
    % Text Node
    \draw (448.3,323) node [anchor=north][inner sep=0.75pt]   [align=left] {${M}$};
    \draw (448.3, 320) -- (448.3, 315);
    \draw (406.8,323) node [anchor=north][inner sep=0.75pt]   [align=left] {${M-1}$};
    \draw (408.3, 320) -- (408.3, 315);

    \draw (375.3,327) node [anchor=north][inner sep=0.75pt]   [align=left] {${\cdots}$};

    \draw (354.3,323) node [anchor=north][inner sep=0.75pt]   [align=left] {${3}$};
    \draw (354.3, 320) -- (354.3, 315);
    \draw (314.3,323) node [anchor=north][inner sep=0.75pt]   [align=left] {${2}$};
    \draw (314.3, 320) -- (314.3, 315);
    \draw (274.3,323) node [anchor=north][inner sep=0.75pt]   [align=left] {${1}$};
    \draw (274.3, 320) -- (274.3, 315);

    % \node at (252, 175.7) [anchor=east] {$Space$};
    \draw (240,83.7) node [anchor=west][inner sep=0.75pt]   [align=left] {${N}$};
    \draw (253,83.7) -- (258, 83.7);
    \draw (220,113.7) node [anchor=west][inner sep=0.75pt]   [align=left] {${N-1}$};
    \draw (253,113.7) -- (258, 113.7);
    \draw (220,143.7) node [anchor=west][inner sep=0.75pt]   [align=left] {${N-2}$};
    \draw (253,143.7) -- (258, 143.7);
    \draw (220,173.7) node [anchor=west][inner sep=0.75pt]   [align=left] {${N-3}$};
    \draw (253,173.7) -- (258, 173.7);

    \draw (245,194.7) node [anchor=north][inner sep=0.75pt]   [align=left] {${\vdots}$};

    \draw (240,243.7) node [anchor=west][inner sep=0.75pt]   [align=left] {${2}$};
    \draw (253,243.7) -- (258, 243.7);
    \draw (240,273.7) node [anchor=west][inner sep=0.75pt]   [align=left] {${1}$};
    \draw (253,273.7) -- (258, 273.7);

    \end{tikzpicture}
    \caption{A surface programmable thermodynamic computer (TC) that implements the rule 110 elementary cellular automata. The vertical direction is the space axis,
    while the horizontal direction is the time axis.
    The computation is done by moving the heat zone (orange colored area) from the input surface (left) to the output surface (right) slowly with a constant speed $v$.
    The TC has a translational invariant structure, and each unit cell (inset) is a $110$ rule cellular automata gadget that consists of 5 mutually correlated spins (white circles).
    % \jinguo{please add symbols for space, time and width}
    \ym{I add symbols. please check if i still need to put original captions here}
    }

    \label{Surface-programmable-thermodynamic computer}
\end{figure}

@@

\subsection{Simulated Annealing}\label{simulated-annealing-algorithm}

@@

\subsection{Cooling computation with an idealized heat bath}\label{cooling-computation-with-moving-heat-bath}

@@

The computational process of the TC is equivalent to finding the ground state of the TC with the spin configuration in the left surface being fixed.

We focus here on the thermodynamics-driven approach to finding ground states, 
specifically the simulated annealing method. This can be seen as the TC interacting with a heat bath, 
as shown in \ref{simulated-annealing-algorithm}, by gradually decreasing the temperature
of the heat bath the TC can be driven to the ground state if the temperature is decreased slowly enough.
However, the cooling process can last for exponential time for some hard spin-glass problems.

We modify the classical simulated annealing (SA) algorithm by introducing a heat zone of width  W , which represents the region of the system that interacts with the heat bath.
To find the ground state of the system, we devided the system into $\lceil\frac{T}{W}\rceil$ parts
where part $i$ consists of the spins in the region $[iW, (i+1)W]$.
However, solving the ground state from the non-deterministic direction is at least as hard as solving the circuit satisfiability problem, which is NP-complete~\cite{Moore2011}.
